---
title: "stacas"
author: "Julia Zhong"
date: today
format:
  html:
    toc: true
    page-layout: full
    embed-resources: true
    code-fold: true
    code-tools: true
    link-external-newwindow: true
lightbox:
  match: auto
  effect: zoom
  desc-position: right
execute:
  warning: false
from: markdown+emoji 
editor_options: 
  chunk_output_type: console
---

## STACAS Data Integration

## Running Code

When you click the **Render** button a document will be generated that includes both content and the output of embedded code. You can embed code like this:

```{r}
library(anndata)
library(dplyr)
library(ggplot2)
library(schard)
library(Seurat)
library(STACAS)

seed <- 1234
set.seed(seed)

data_path <- "/Users/juliazhong/Documents/graduate school/20.440/project"
```

take a look at the .h5ad file as an AnnData object
```{r}
ad <- read_h5ad(paste0(data_path, "/adata_2025-04-20.h5ad"))
ad
```

convert AnnData object to Seurat object
```{r}
# IMPORTANT! some other methods result in the loss of the raw counts matrix during the conversion. in Python, need to run:
# adata.raw = adata.X.copy() when .X is the raw matrix, AND
# use this schard function with use.raw=TRUE
object = schard::h5ad2seurat(paste0(data_path, "/adata_2025-04-20.h5ad"), use.raw=TRUE)
```

```{r}
nfeatures <- 1000
ndim <- 20
object <- FindVariableFeatures(object, nfeatures = nfeatures) %>%
  NormalizeData() %>% ScaleData() %>%
  RunPCA(npcs=ndim) %>% RunUMAP(dims=1:ndim)
```

```{r}
pre <- DimPlot(object, group.by = "sample") + theme(aspect.ratio = 1) +
  ggtitle("Dataset prior to STACAS integration")

pre
```

```{r}
object_integrated <- object %>% SplitObject(split.by = "sample") %>%
      Run.STACAS(dims = 1:ndim, anchor.features = nfeatures) %>%
      RunUMAP(dims = 1:ndim) 
```

```{r}
post <- DimPlot(object_integrated, group.by = "sample") + theme(aspect.ratio = 1) +
  ggtitle("Dataset after STACAS integration")

pre | post
```

```{r}
DimPlot(
  object,
  group.by = "sample",
  split.by = "sample",
  ncol = 4
) + 
  theme(aspect.ratio = 1) +
  ggtitle("Dataset prior to STACAS integration") +
  NoLegend()
```

```{r}
DimPlot(
  object_integrated, 
  group.by = "sample", 
  split.by = "sample",
  ncol = 4
) + theme(aspect.ratio = 1) +
  ggtitle("Dataset after STACAS integration") +
  NoLegend()
```

convert post-data integration Seurat object back to AnnData for further analysis in Python
```{r}
sceasy::convertFormat(object_integrated, from="seurat", to="anndata", outFile=paste("../integrated_", Sys.Date(), ".h5ad", sep=""))
```

